---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# reproducibleai

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/MVR-GIS/reproducibleai/graph/badge.svg)](https://app.codecov.io/gh/MVR-GIS/reproducibleai)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle)
[![License: CC0](https://img.shields.io/badge/license-CC0-blue.svg)](http://choosealicense.com/licenses/cc0-1.0/)
<!-- badges: end -->

`{reproducibleai}` creates, migrates, and validates durable repository context
for reproducible AI-assisted development. Instead of copying chat instructions
and preserving transcripts as working context, it creates a concise root
`AGENTS.md` that routes work into maintained artifacts under `dev/`.

The standard separates standing instructions from goals, architecture,
decisions, governance, workflows, schemas, feature specifications, and concise
checkpoints. Git preserves superseded history. The package can also measure how
effectively Codex follows those routes across repeated isolated runs.

## Installation

```{r eval=FALSE}
# install.packages("pak")
pak::pak("MVR-GIS/reproducibleai")
```

The core package is local-first. Scaffolding, migration planning and
application, structural validation, competency-question authoring, offline
scoring, summaries, and report rendering do not require a Codex installation,
cloud SaaS account, or administrator rights.

Live agentic-routing evaluation is an optional connected capability. It
requires:

- a separately installed Codex CLI;
- an authenticated Codex CLI session;
- outbound connectivity to the Codex service; and
- permission under the computer's application, endpoint, identity, and network
  policies.

On Windows, OpenAI's standalone installer installs into the user profile and
does not require npm:

```powershell
powershell -ExecutionPolicy ByPass -c "irm https://chatgpt.com/codex/install.ps1 | iex"
```

Restart the R IDE after installation, then run:

```{r eval=FALSE}
check_agentic_routing_prerequisites(path = ".")
```

The diagnostic also discovers the standard user-local Windows installation
when an IDE has a stale `PATH`. It does not install software, authenticate,
contact a model, or consume model usage. On managed or disconnected equipment,
live execution may remain prohibited even when package installation succeeds.
`{reproducibleai}` reports that limitation and retains its offline
functionality; it never attempts to bypass organizational controls.

## Scaffold a new R package

```{r eval=FALSE}
library(reproducibleai)

use_agentic_context(
  path = ".",
  profiles = c("base", "r-package")
)

validate_agentic_context(".")
```

Scaffolding is deterministic and idempotent. Existing repository-owned content
is never overwritten.

## Migrate an existing repository

```{r eval=FALSE}
plan <- plan_agentic_context_migration(".")

# Review the structured operations before changing the repository.
plan$operations

result <- apply_agentic_context_migration(
  plan,
  approved = TRUE
)
```

Planning is read-only: legacy content remains in place. Application creates and
validates replacements before removing superseded files. Destructive operations
are refused when a source is untracked, has uncommitted changes, or requires
manual review.

## Profiles

Standard 0.1 includes:

- `base`, which applies to every repository
- `r-package`, which adds routes and workflow guidance for R package development

`detect_agentic_context_profiles()` reports evidence for candidate profiles but
does not silently select or apply them.

## Scope

The agentic-routing evaluation API defines private competency questions, runs
fresh read-only `codex exec` sessions, scores evidence routing and expected
answer content, and writes aggregate health reports.

```{r eval=FALSE}
check_agentic_routing_prerequisites(path = ".")

benchmark <- derive_agentic_routing_questions(".")
print(benchmark) # every generated question begins as pending

ids <- vapply(benchmark$questions, function(x) x$id, character(1))
benchmark <- review_agentic_routing_benchmark(
  benchmark,
  approve = ids # approve only after reading every candidate
)

evaluation <- run_agentic_routing_evaluation(
  path = ".",
  questions = benchmark,
  repetitions = 5,
  approved = TRUE
)

health <- summarize_agentic_routing(evaluation)
write_agentic_routing_report(
  health,
  "dev/governance/agentic-routing-health.md"
)
```

Keep private fixtures and raw runs outside the evaluated repository. Freeze
reviewed generated benchmarks outside the repository with
`write_agentic_routing_benchmark()` so specification variants use the same
criterion. Model responses remain stochastic; health reports summarize repeated
observations rather than claiming deterministic model behavior. Live evaluation
is intended as an explicitly initiated local development diagnostic, not as an
automatic package-build or CI step. A committed aggregate report can communicate
the observed health of repository context to maintainers and end-users.

## Bug reports

Please open an issue at
[https://github.com/MVR-GIS/reproducibleai/issues](https://github.com/MVR-GIS/reproducibleai/issues).
