---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# reproducibleai

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/MVR-GIS/reproducibleai/graph/badge.svg)](https://app.codecov.io/gh/MVR-GIS/reproducibleai)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle)
[![License: CC0](https://img.shields.io/badge/license-CC0-blue.svg)](http://choosealicense.com/licenses/cc0-1.0/)
<!-- badges: end -->

`{reproducibleai}` creates, migrates, and validates durable repository context
for reproducible AI-assisted development. Instead of copying chat instructions
and preserving transcripts as working context, it creates a concise root
`AGENTS.md` that routes work into maintained artifacts under `dev/`.

The standard separates standing instructions from goals, architecture,
decisions, governance, workflows, schemas, feature specifications, and concise
checkpoints. Git preserves superseded history.

## Installation

```{r eval=FALSE}
# install.packages("pak")
pak::pak("MVR-GIS/reproducibleai")
```

## Scaffold a new R package

```{r eval=FALSE}
library(reproducibleai)

use_agentic_context(
  path = ".",
  profiles = c("base", "r-package")
)

validate_agentic_context(".")
```

Scaffolding is deterministic and idempotent. Existing repository-owned content
is never overwritten.

## Migrate an existing repository

```{r eval=FALSE}
plan <- plan_agentic_context_migration(".")

# Review the structured operations before changing the repository.
plan$operations

result <- apply_agentic_context_migration(
  plan,
  approved = TRUE
)
```

Planning is read-only: legacy content remains in place. Application creates and
validates replacements before removing superseded files. Destructive operations
are refused when a source is untracked, has uncommitted changes, or requires
manual review.

## Profiles

Standard 0.1 includes:

- `base`, which applies to every repository
- `r-package`, which adds routes and workflow guidance for R package development

`detect_agentic_context_profiles()` reports evidence for candidate profiles but
does not silently select or apply them.

## Scope

This release establishes deterministic context scaffolding and migration. Model
execution, competency-question scoring, routing sensitivity analysis, health
reports, and prompt tuning are intentionally reserved for a later evaluation
feature.

## Bug reports

Please open an issue at
[https://github.com/MVR-GIS/reproducibleai/issues](https://github.com/MVR-GIS/reproducibleai/issues).
