---
title: "Durable agentic context for a repository"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Durable agentic context for a repository}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Why durable context?

An AI-assisted development session is most useful when the agent can quickly
find current goals, architecture, decisions, contracts, and workflows. Full
transcripts and copied instruction manuals consume context while allowing
important facts to drift.

The `{reproducibleai}` standard uses:

- a concise root `AGENTS.md` for standing rules and routes,
- purpose-specific maintained artifacts under `dev/`,
- checkpoints only for useful resumable state, and
- Git for superseded history.

## New repositories

```{r eval=FALSE}
use_agentic_context(
  path = ".",
  profiles = c("base", "r-package")
)
```

The function preflights every target. It creates missing files, skips identical
files, and refuses to overwrite differing repository-owned content.

## Existing repositories

Migration is intentionally two-phase:

```{r eval=FALSE}
plan <- plan_agentic_context_migration(".")
plan$operations

result <- apply_agentic_context_migration(plan, approved = TRUE)
```

The plan is a durable, inspectable crosswalk. Application is blocked by any
manual-review operation. When approved, it creates replacements, copies mapped
legacy artifacts into their new routes, validates the structure, and removes
superseded tracked sources last.

## Validation

```{r eval=FALSE}
report <- validate_agentic_context(".")
report$valid
report$findings

# Suitable for CI:
validate_agentic_context(".", strict = TRUE)
```

Validation checks structure, required routes, the installed standard version,
and changes relative to seeded file hashes. Local edits are warnings rather than
errors because scaffolded files become repository-owned after creation.
