---
title: "Evaluate agentic routing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Evaluate agentic routing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Agentic-context validation confirms that a repository has the expected
structure. Routing evaluation asks a different question: when Codex receives a
task, does it find the right maintained evidence and produce the expected
answer efficiently?

The harness is deterministic, but model responses are stochastic. Interpret
results as repeated-run distributions rather than byte-identical tests.

## Capability and environment preflight

Most `{reproducibleai}` functionality is available without Codex, a cloud
account, or administrator rights. Live routing evaluation is different: it
starts authenticated model runs and therefore requires a separately installed
Codex CLI, connectivity, account entitlement, and organizational permission.

On Windows, the standalone installer uses a user-local installation and does
not require npm:

```powershell
powershell -ExecutionPolicy ByPass -c "irm https://chatgpt.com/codex/install.ps1 | iex"
```

Restart the R IDE after installing. Then perform a no-usage preflight:

```{r eval=FALSE}
library(reproducibleai)

check_agentic_routing_prerequisites(
  path = "C:/workspace/FluvialGeomorph/fluvgeo"
)
```

Automatic discovery checks `PATH` and the standard user-local Windows
installation. Supply `codex = "C:/path/to/codex.exe"` only for a nonstandard
installation. The diagnostic checks the executable, saved authentication,
agentic context, and Git state without contacting a model. Network and
enterprise execution policy remain untested until a live run begins.

On managed government-furnished, disconnected, or no-admin equipment, package
installation and offline capabilities remain useful even if application
allow-listing, identity policy, or network controls prohibit live Codex
evaluation. Do not bypass those controls. Prepare or inspect fixtures and
reports offline, and perform live evaluation only in an approved environment.

## Define private competency questions

```{r eval=FALSE}
library(reproducibleai)

questions <- list(
  new_agentic_routing_question(
    id = "current-priority",
    prompt = "What is the current development priority and its exit criteria?",
    required_paths = "dev/goals/project-plan.md",
    allowed_paths = c("AGENTS.md", "dev/architecture/design.md"),
    expected_terms = c("current objective", "exit criteria"),
    forbidden_terms = "session transcript"
  ),
  new_agentic_routing_question(
    id = "architecture-boundary",
    prompt = "Which repository owns this behavior and what evidence supports that?",
    required_paths = c(
      "AGENTS.md",
      "dev/architecture/backend-ecosystem.md"
    ),
    allowed_paths = "dev/workflows/backend-change-assessment.md",
    expected_terms = c("ownership", "downstream")
  )
)

fixture <- file.path(tempdir(), "fluvgeo-routing-questions.json")
write_agentic_routing_questions(questions, fixture)
questions <- read_agentic_routing_questions(fixture)
```

Keep the fixture outside the repository under evaluation. Otherwise Codex can
inspect the private rubric.

## Run repeated isolated sessions

```{r eval=FALSE}
evaluation <- run_agentic_routing_evaluation(
  path = "C:/workspace/FluvialGeomorph/fluvgeo",
  questions = questions,
  repetitions = 5,
  approved = TRUE,
  model = NULL
)
```

Each repetition uses a fresh `codex exec` session with:

- `--ephemeral`;
- `--sandbox read-only`;
- `--ask-for-approval never`;
- `--ignore-user-config`;
- `--json`; and
- `--output-schema`.

The function uses existing Codex authentication. It does not accept or record
API keys. Raw JSONL, responses, and stderr remain in an external temporary
directory unless another external directory is supplied.

`codex exec` and structured output are documented in OpenAI's
[non-interactive mode](https://learn.chatgpt.com/docs/non-interactive-mode) and
[CLI reference](https://learn.chatgpt.com/docs/developer-commands?surface=cli#cli-codex-exec).

## Summarize and report

```{r eval=FALSE}
health <- summarize_agentic_routing(evaluation)
health$health_score
health$by_question

write_agentic_routing_report(
  health,
  path = file.path(
    "C:/workspace/FluvialGeomorph/fluvgeo",
    "dev",
    "governance",
    "agentic-routing-health.md"
  )
)
```

The durable report contains only aggregate metrics and recommendations. Inspect
external raw runs before changing instructions, and validate proposed
improvements on held-out questions or repositories.

Run this workflow deliberately during local development when agentic context
changes. Do not place live evaluation in package builds or CI. Commit the
aggregate report as a transparent quality and troubleshooting artifact; never
commit the private fixtures or raw traces.
